Forecasting the 2024 U.S. Presidential Election: Kamala Harris' Projected Victory

### 2024-11-03 paper/paper.qmd
committing text changes to paper.qmd

# Data {#sec-data}
## Overview
This paper utilizes the dataset of the 2024 Presidential Election cycle, obtained from [@FiveThirtyEight]. The dataset is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results. It consists of polling information from various pollsters such as the New York Times, ActiVote, Morning Consult, etc. For every pollster, relevant variables such as state, polling date, transparency score, candidate name, and percentage of support for the appropriate candidate are included. The dataset compiles data from all major pollsters in the US and provides us with an understanding of the current state of the presidential election.

The present analysis focuses on Donald Trump and Kamala Harris, the two leading candidates in the current presidential election race. For all the following variables mentioned, we collected observations for both of the candidates. We collect observations from the dataset where the ‘candidate_name’ is Trump and Harris. We are interested in the support for both candidates over time, for each state, therefore, we collect observations of ‘end_date’ and ‘state’. Furthermore, since there are many pollsters present in the dataset, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high ‘numeric_grade’ and ‘transparency_score’, details of data quality are further explained in {additional data details}. The electorate support for each candidate is denoted as ‘pct’.

All of the variables mentioned above are used during the steps of data cleaning, simulation, testing, and analysis. We use the statistical programming language R [cite R]... to perform data simulation, testing, cleaning, and analysis.

@@

MISSING - explain table

>>>>

# Data {#sec-data}
## Overview
NEEDS EDIT & ELABORATE - This paper utilizes the dataset of the 2024 Presidential Election cycle, obtained from [@FiveThirtyEight]. The dataset is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results. It consists of polling information from various pollsters such as the New York Times, ActiVote, Morning Consult, etc. For every pollster, relevant variables such as state, polling date, transparency score, candidate name, and percentage of support for the appropriate candidate are included. The dataset compiles data from all major pollsters in the US and provides us with an understanding of the current state of the presidential election.

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-stateaverage
#| fig-cap: "Table displaying the percentage of support for Donald Trump and Kamala Harris across key battleground states in the lead-up to the 2024 United States Presidential Election. The table lists states with their respective support percentages, highlighting the competitive landscape as voters prepare to make their decisions on Election Day. Notably, Trump maintains a slight lead in states such as Arizona and Georgia, while Harris shows stronger support in states like Michigan and Nevada. This data provides insight into the current polling dynamics and potential electoral outcomes in these critical states."

# Aggregating polling percentages for Trump and Harris by state
state_average <- clean_poll_data %>%
   group_by(state) %>%
   summarise(
     Trump_pct = mean(Trump_pct, na.rm = TRUE),
     Harris_pct = mean(Harris_pct, na.rm = TRUE)
   )

 # Prepare table for display with rounding to 1 decimal place
 average_table <- state_average |>
   mutate(
     Trump_pct = round(Trump_pct, 1),
     Harris_pct = round(Harris_pct, 1)
   ) |>
   select(
     State = state, 
     "Trump %" = Trump_pct, 
     "Harris %" = Harris_pct, 
   )

 # Render the table
 knitr::kable(average_table)
```
